### 2020-02-09 Rcodes/JAGS_practice.Rmd
jags lagged: errors to be questioned

---
title: "R Notebook"
output: html_notebook
---
# Following an online tutorial (no data...)

Context:Normal model with unknown mean and variance.

summary: 
in the model, as a string, we put the likelihood and the parameters of interest with their prior distribution.
```{r}
install.packages('runjags')
install.packages('coda')
```

```{r}
library(runjags)
modelString = "
model{
for (i in 1:N) {
y[i] ~ dnorm(mu, phi)
} 
mu ~ dnorm(mu_0, phi_0)
phi ~ dgamma(alpha, beta)
}
"

```
- note that they take precison for Normal distribution
-use 'd' to describe density: 'r' never used.

From here, the list named as the_data, we plug in the values of the parameters for the prior including the number of observation for likelihood (N).

Posterior result is obtained by run.jags function where we plug in our data (the_data) which we set the values of the parameters for prior, we state our interest parameters as monitor. I assume modelString goes into the first argument here.
```{r}
y <- CEdata$LogTotalExpLastQ
N <- length(y)
the_data <- list(
  'y'=y,
  'N'=N,
  'mu_0'=5,
  'phi_0'=1/1^2,
  'alpha'=1,
  'beta'=1) # this we where the pre-set priors given to the model
  posterior <- run.jags(
    'norm_model.jags',
    data = the_data,
    monitor = c('mu','phi'),
    n.chains=1,
    adapt=1000,
    burnin=2000,
    sample=5000,
    thin=1
  )

summary(posterior)

```

# Another JAGS Tutorial 
[from here](https://rstudio-pubs-static.s3.amazonaws.com/272658_ae4d482c86514674be17042c852ebbfc.html)

```{r}
setwd("../../meta-analysis-Bayesian/")
data =read.csv('data1.csv')
head(data)
```

```{r Defining values}
library(coda)
N <- length(data$y)

data <- list('N'=N, 
             'y'=data$y,
             'V'=data$V)

inits <- list(d=0.0)

```

```{r Enter the Model}
cat("model{
for (i in 1:N) {
P[i] <- 1/V[i]
y[i]~dnorm(d,P[i])
}
# Define the priors
d ~ dnorm(0,0.00001)

# Transfomr the ln(OR) to OR
OR <- exp(d)
}

    ", file='model_test.txt')
```

```{r}
library(rjags)
jags.m <- jags.model(file='model_test.txt',data=data, inits=inits, n.chains = 1, n.adapt = 500,)
```
```{r}
params <- c('d','OR')
samples <- coda.samples(jags.m, params, n.iter=10000)
```

```{r}
summary(samples)
```

```{r}
summary(window(samples, start=5001))
```

```{r}
plot(samples)
```

_______________________________________________________________

# Opioid Overdose Modeling 
First, we bring the two data sets (survey data and ambulance attended overdoses data) into here. It is done by manual labor for now but it should be automatical when the pipeline with pymc3 code is built (probably with reticulate package).
```{r}
df <- data.frame(
  o_t= c(2475,  262,  318,  149, 1151,   39, 2997,  242,  702,  377, 2088,76), 
  u_t=c(1969,  217,  253,  119,  934,   34, 2392,  196,  569,  308, 1655,55), 
  x_a= c(799, 798, 795, 816, 805, 794, 793, 780, 773, 779, 788, 813)
)
n_a=1000
N= 10000
```

runjags package

Todo:

1. Import relavent functions from the pymc3 Jupyter notebook.
2. Obtain the same data sets.

To do 1, I need to save the functions into a module or a library.
```{r Package preparing}
# package connecting R and Python
install.packages("reticulate", dependencies = TRUE)
# JAGS
install.packages('rjags')
library(reticulate)
library(rjags)
```
```{r}
py_install("pandas")
py_run_file("../report_revised_2.py")
```

```{r}
install.packages('RSQLite')
library(RSQLite)
```
```{r}
conn <- dbConnect(RSQLite::SQLite(), "notebook_env.db")
alltables = dbListTables(conn)
```

```{r}
install.packages('ProjectTemplate')
library('ProjectTemplate')
```

```{r}
db.reader('notebook_env.db', '../notebook_env.db', 'data')
```