### 2020-03-04 Rcodes/JAGS_overdoses.Rmd
almost done with JAGS and flu shot is important

---
title: "JAGS Model"
author: "Tom Park"
date: '2020-02-08'
output:
  pdf_document: default
  html_document: default
---
# Basic Model: Ambulance + Overdose
## Ambulance Call-outs Model
$n_{A}$: sample size   
$x_{A}$: the total number who confirmed they did call an ambulance  
$p_{A}$: probability of a person call an ambulance   
$$x_{A} \sim Bin(n_{A},p_{A})$$
**We assume $n_{A}=1000, p_A = 0.8$.**  
**Suppose the prior of $p_A$ is noninformative.** 
$$p(p_A) \sim Beta(1,1)$$

## Overdose Model  
Now we plug in this values into the overdose model and obtain possible $O_t$ values **assuming we have $U_t$ values.**

Also, we have priors.
$$z_{t} \sim N(\mu, \sigma^{2})$$
$$\lambda_{t}^{OD} = \exp(z_{t})$$
$$O_{t} \sim Poi(\lambda_{t}^{OD}N)$$
$$U_t \sim Bin(O_t, p_A)$$
For simplicity we set N =10000 for now. We need to generate reasonable $U_t$ values first. Note that $U_t$ comes from $\mu, \sigma$ following all the way through the overdose model.  
$\mu=\log0.05, \sigma=1, N=10000$.   
We suppose survey data exists: ($n_A, x_A$) known.

**We set for our prior parameters:** $$\mu \sim U(-10,0)$$  $$\sigma \sim U(0,5)$$

```{r message=FALSE, warning=FALSE}
# install packages
if (!require(rjags)) install.packages("rjags", dependencies = TRUE)
if (!require(coda)) install.packages("coda", dependencies = TRUE)
if (!require(tidyverse)) install.packages("tidyverse", dependencies = TRUE)
if (!require(tinytex)) install.packages("tinytex", dependencies = TRUE)

library('rjags')
library('coda')
library('tidyverse')
library('tinytex')

```
The data is the same data from pymc3 with Python.  
Todo: build a pipeline to connect the python (pymc3) and R (JAGS)

```{r}
df <- read.csv('./basic_data.csv')
df$X <- NULL 
head(df)
```
Now we set the model which defines the relations of overdose model and ambulance call model.

The model defined as follows.
```{r Enter the models}
cat("model{
## define the priors

p_a ~ dbeta(alpha, beta)
mu_z ~ dunif(mu_a, mu_b)
sigma_z ~ dunif(sigma_a, sigma_b)

 ## ambulance model
for (i in 1:n) {
  #Likelihood
  x_a[i] ~ dbin(p_a, n_a) # each survey result for month

}

# overdose 
for (i in 1:n) {

  ## the latent variables
  z_t[i]~ dnorm(mu_z, 1/(sigma_z^2))
  lmb_t[i] <- exp(z_t[i])

  ## overdose model
  o_t[i] ~ dpois(lmb_t[i]*N) # total overdoses per month
  # Note that from pymc3 gamma was used instead of Pois dist
  u_t[i] ~ dbin(p_a, o_t[i]) # ambulanced overdoses per month
}

}", file='basic_model.txt')
```

Pre-set variables.
```{r Defining Values}
n_T <-  length(df$o_t) # we have 12 samples
n_a <- 1000
N <- 10000
u_t <- df$u_t
x_a <- df$x_a
```

Define the list providing the values of the variables and the parameters for the priors of the model. 
```{r from variable name to \'variable_name\'}
dat <- list(
            # priors for ambulance model
            'alpha' = 1,
            'beta' =  1,

            # priors for overdose model 
            'mu_a'=(-10),
            'mu_b'=0,
            'sigma_a'=0,
            'sigma_b'=5,

            # likelihood 
            'u_t'=u_t, # giving data 
            'x_a'=x_a,   # giving data
            'N'=N, # the population 10000
            'n'= n_T, # total months 12
            'n_a'=n_a # survey size 1000

            )
```
**Note: for the list object usually named 'data' or 'dat' in JAGS context, do not use arrow but use equal sign to define elements of the list.**
``` {r run the first model}
chains=2
# inits = list()
simple.model <- jags.model(file='basic_model.txt',
                          data=dat,
                          n.chains = chains)
simple.model
```

### O_t

```{r}
params= c('o_t','p_a')
samples <- coda.samples(simple.model, params, n.iter = 1000)
# guess it's getting posterior samples ?
```

```{r }
interations = 1000
burnin= floor(interations/2)
summary(window(samples), start=burnin)
```

```{r include=FALSE}
plot(samples)
```
# q1: what is the equivalent plot that we can see we have enough iteration?

### Boxplots of O_t

# q2: I see two elements from the samples list. Which one I should use it or should I use both?

```{r yield a data frame}

pst_mtx = as.matrix(samples)
temp = pst_mtx[,1:12]
p_a <- pst_mtx[,13]
head(temp)
length(p_a)
colnames(temp) <- seq(1,12)
df_o_t <- as.data.frame(temp)
head(df_o_t)
str(df_o_t)
```

```{r tidy the data frame}
trace_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
str(trace_o_t)
head(trace_o_t,n = 24)
```

```{r}
# real values of the data set from pymc3 samples
o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
# factorizing the months for box plot visualization
o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))

ggplot()+
  # boxplot from the trace 
  geom_boxplot(aes(x=month,y=o_t), color='grey',data = trace_o_t)+
  # real values as red dots
  geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')

```

## Predictive Posterior Checks

```{r}
df_u_t <- data.frame()

for (i in 1:m) {
  obs <- rbinom(n=12,size = as.numeric(df_o_t[i,]),prob =p_a[i])
  df_u_t=rbind(df_u_t,obs)
}
colnames(df_u_t) <- factor(seq(1,12),levels = seq(1,12))
str(df_u_t)
```

### $U_t$: Predictive Posterior Checks

```{r}
ppc_u_t <- gather(df_u_t, key = 'month',value = 'u_t')
summary(ppc_u_t)
u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
# u_t_values$month <- factor(u_t_values$month,levels = seq(1,12))
ppc_u_t$month <- factor(ppc_u_t$month,levels = seq(1,12))

ggplot()+geom_boxplot(aes(x=month,y=u_t),data = ppc_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')
```

# todo: finish this part.
### $x_A$: Predictive Posterior Checks

```{r Generate posterior predictive samples of $x_a$}

df_x_a <- vector()
for (i in 1:n_T) {
  obs <- rbinom(m,n_a,p_a)
  df_x_a <- cbind(df_x_a,obs)
}

head(df_x_a)
colnames(df_x_a) <- seq(1:12)
```

```{r}

df_x_a <- as.data.frame(df_x_a)
ppc_x_a <- gather(df_x_a, key = 'month',value = 'x_a')

ppc_x_a$month <- factor(ppc_x_a$month,levels = seq(1,12))

x_a_values=data.frame('month'=seq(1,12),'x_a'=df$x_a)
x_a_values$month <- factor(x_a_values$month,levels = seq(1,12))

ggplot()+geom_boxplot(aes(x=month,y=x_a),data = ppc_x_a)+geom_point(aes(x=x_a_values$month, y=x_a_values$x_a),color='red')
```

# Contamination of $p_A$

Now, suppose the survey data gives us a wrong (biased) $p_A$ value.

Bias = $\theta -\hat{\theta} = p_A- \hat{p}_A$   
$\hat{p}_A = p_A + bias(p_A)$

Three more data sets are given: unbiased, overestimated, underestimated $p_A$.

```{r}
## write a function that led to compare o_t, u_t and x_a. 
#first we need a fuction that gives us data, model, trace, and ppc.
test_robust <- function(file=file, random=1, N=10000, p_a=0.8, bias = -0.2, n_a=1000, n_T=12) {
  df <- read.csv(file = file)
  df$X <- NULL 
  df$month <- seq(1:12)
# obtain the (biased) data 
  dat <- list(
            # priors for ambulance model
            'alpha' = 1,
            'beta' =  1,

            # priors for overdose model 
            'mu_a'=(-10),
            'mu_b'=0,
            'sigma_a'=0,
            'sigma_b'=5,

            # likelihood 
            'u_t'=df$u_t, # giving data 
            'x_a'=df$x_a,   # giving data
            'N'=N, # the population 10000
            'n'= n, # total months 12
            'n_a'=n_a # survey size 1000

            )
# run the model 
  chains=2
  # target 1 to save
  simple.model <- jags.model(file='basic_model.txt',
                          data=dat,
                          n.chains = chains)

# get the samples of O_t, p_a
  params= c('o_t','p_a')
  # target 2 to save
  samples <- coda.samples(simple.model, params, n.iter = 1000)

  # tidy p_a: target 3 
  p_a <- pst_mtx[,13]

  ## tidy o_t
  pst_mtx = as.matrix(samples)
  temp = pst_mtx[,1:12]
  colnames(temp) <- seq(1,12)
  df_o_t <- as.data.frame(temp)
  trace_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
  trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
  # real values of the data set from pymc3 samples
  # o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
  # # factorizing the months for box plot visualization
  # o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))
  # boxplots of o_t

  # tidy u_t pp samples
  df_u_t <- data.frame()
  for (i in 1:m) {
   obs <- rbinom(n=12,size = as.numeric(df_o_t[i,]),prob =p_a[i])
   df_u_t=rbind(df_u_t,obs)
  }
  colnames(df_u_t) <- factor(seq(1,12),levels = seq(1,12))
  ppc_u_t <- gather(df_u_t, key = 'month',value = 'u_t')
  ppc_u_t$month <- factor(ppc_u_t$month,levels = seq(1,12))
  # u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)

  ## tidy x_a
  df_x_a <- vector() 
  for (i in 1:n) {
    obs <- rbinom(m,n_a,p_a)
    df_x_a <- cbind(df_x_a,obs)
  }
  colnames(df_x_a) <- seq(1:12)
  df_x_a <- as.data.frame(df_x_a)
  ppc_x_a <- gather(df_x_a, key = 'month',value = 'x_a')
  ppc_x_a$month <- factor(ppc_x_a$month,levels = seq(1,12))
  # x_a_values=data.frame('month'=seq(1,12),'x_a'=df$x_a)
  # x_a_values$month <- factor(x_a_values$month,levels = seq(1,12))
  ppc = list('u_t'=ppc_u_t, 'x_a'=ppc_x_a)
  mylist=list('data'=df,'model'=simple.model,'trace'=trace_o_t,'ppc'=ppc)
  return (mylist)
}

# then visualization function which gives us three different types of boxplots.
```

```{r}
my_list_unbiased = test_robust(file = './basic_data.csv', bias =0)
my_list_under = test_robust(file = './under_p_a_data.csv',bias=-0.2)
my_list_over = test_robust(file='./over_p_a_data.csv',bias = +0.1)
```

```{r Check whether we use the same data set}
print(my_list_unbiased$data$u_t)
print(my_list_under$data$u_t)
print(my_list_over$data$u_t)

print(my_list_unbiased$data$x_a)
print(my_list_under$data$x_a)
print(my_list_over$data$x_a)
```
```{r}
 head(my_list_unbiased$ppc$u_t)
```

```{r}
# finish this function
visualization <- function(mylist= None, post= F, u_t = F, x_a = F, string='string') {
  data= mylist$data
  if (post == T) {
    # boxplots of o_t
    trace_o_t = mylist$trace
  p <- ggplot()+
  # boxplot from the trace 
  geom_boxplot(aes(x=month,y=o_t), color='grey',data = trace_o_t)+
  # real values as red dots
  geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')

  }

  if (u_t == T) {
    ppc_u_t = mylist$ppc$u_t
    # boxplots of u_t
  p <- ggplot()+geom_boxplot(aes(x=month,y=u_t),data = ppc_u_t)+geom_point(aes(x=data$month, y=data$u_t),color='red')

  }

  if (x_a == T) {
    p <- ppc_x_a = mylist$ppc$x_a
    ggplot()+geom_boxplot(aes(x=month,y=x_a),data = ppc_x_a)+geom_point(aes(x=data$month, y=data$x_a),color='red')
  }
  return(p)
}
```

```{r}
visualization(my_list_unbiased, post= T)
```

# Trim here and make a function and give us the plots.
```{r}
## Trim here and make a function and give us the plots.
df <- read.csv('./basic_data.csv')
df$X <- NULL 
# obtain the (biased) data 
dat <- list(
            # priors for ambulance model
            'alpha' = 1,
            'beta' =  1,

            # priors for overdose model 
            'mu_a'=(-10),
            'mu_b'=0,
            'sigma_a'=0,
            'sigma_b'=5,

            # likelihood 
            'u_t'=df$u_t, # giving data 
            'x_a'=df$x_a,   # giving data
            'N'=N, # the population 10000
            'n'= n, # total months 12
            'n_a'=n_a # survey size 1000

            )
# run the model 
chains=2
simple.model <- jags.model(file='basic_model.txt',
                          data=dat,
                          n.chains = chains)

# get the samples of O_t, p_a
params= c('o_t','p_a')
samples <- coda.samples(simple.model, params, n.iter = 1000)

# tidy p_a
p_a <- pst_mtx[,13]

## tidy o_t
pst_mtx = as.matrix(samples)
temp = pst_mtx[,1:12]
colnames(temp) <- seq(1,12)
df_o_t <- as.data.frame(temp)
trace_o_t <- gather(df_o_t, key = 'month',value = 'o_t')
trace_o_t$month <- factor(trace_o_t$month,levels = seq(1,12))
# real values of the data set from pymc3 samples
o_t_values=data.frame('month'=seq(1,12),'o_t'=df$o_t)
# factorizing the months for box plot visualization
o_t_values$month <- factor(o_t_values$month,levels = seq(1,12))
# boxplots of o_t
ggplot()+
  # boxplot from the trace 
  geom_boxplot(aes(x=month,y=o_t), color='grey',data = trace_o_t)+
  # real values as red dots
  geom_point(aes(x=o_t_values$month, y=o_t_values$o_t),color='red')

# tidy u_t pp samples
df_u_t <- data.frame()
for (i in 1:m) {
  obs <- rbinom(n=12,size = as.numeric(df_o_t[i,]),prob =p_a[i])
  df_u_t=rbind(df_u_t,obs)
}
colnames(df_u_t) <- factor(seq(1,12),levels = seq(1,12))
ppc_u_t <- gather(df_u_t, key = 'month',value = 'u_t')
ppc_u_t$month <- factor(ppc_u_t$month,levels = seq(1,12))
u_t_values=data.frame('month'=seq(1,12),'u_t'=df$u_t)
# boxplots of u_t
ggplot()+geom_boxplot(aes(x=month,y=u_t),data = ppc_u_t)+geom_point(aes(x=u_t_values$month, y=u_t_values$u_t),color='red')

## tidy x_a
df_x_a <- vector() 
for (i in 1:n) {
  obs <- rbinom(m,n_a,p_a)
  df_x_a <- cbind(df_x_a,obs)
}
colnames(df_x_a) <- seq(1:12)
df_x_a <- as.data.frame(df_x_a)
ppc_x_a <- gather(df_x_a, key = 'month',value = 'x_a')
ppc_x_a$month <- factor(ppc_x_a$month,levels = seq(1,12))
x_a_values=data.frame('month'=seq(1,12),'x_a'=df$x_a)
x_a_values$month <- factor(x_a_values$month,levels = seq(1,12))

ggplot()+geom_boxplot(aes(x=month,y=x_a),data = ppc_x_a)+geom_point(aes(x=x_a_values$month, y=x_a_values$x_a),color='red')
```

## Reference 
[first tutorial](https://www.youtube.com/watch?v=F9jxvhXc_1w)  
[second tutorial](https://rstudio-pubs-static.s3.amazonaws.com/272658_ae4d482c86514674be17042c852ebbfc.html)   
[JAGS manual](https://web.sgh.waw.pl/~atoroj/ekonometria_bayesowska/jags_user_manual.pdf)   
[error handling guide](https://www4.stat.ncsu.edu/~reich/ABA/code/errors)